Predictions for detecting BHNS and other double compact objects binaries with LISA

### 2021-04-10 paper/tex/sections/methods.tex
reorganise results: table to appendix, sc plot sooner
add LEGWORK details to methods

\subsubsection{Population Overview}

@@

\subsubsection{Initial Conditions}

@@

\subsection{Galaxy Synthesis}

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=\columnwidth]{galaxy_metallicity.png}
%     \caption{A random instance of the Milky Way of $10^5$ binaries, created using distributions from \citet{Frankel+2018}. The colour of each point represents its metallicity (with $Z \in [0.0001, 0.022]$). The top panel shows a side-on view and the bottom panels shows a view from above.}
%     \label{fig:galaxy_metallicity}
% \end{figure}

@@

\subsection{Gravitational Wave Inspiral}

@@

\subsection{Detectability}

We define a binary as detectable if its gravitational wave signal has a signal-to-noise ratio of greater than 7 \citep[e.g.][]{Breivik+2020, Korol+2020}. The sky-, polarisation- and orientation-averaged signal-to-noise ratio, $\rho$, of an inspiraling binary can be calculated with the following\footnote{For full derivations of the equations in this section, please see Appendix \ref{app:strain_snr_deriv}.} \citep[e.g.][]{Finn+2000}
\begin{equation}\label{eq:snr}
    \rho^2 = \sum_{n=1}^{\infty} \int_{f_{n, i}}^{f_{n, f}} \frac{h_{c, n}^{2}}{f_{n}^{2} S_{\rm n}\left(f_{n}\right)} \dd{f_n},
\end{equation}
where $n$ is a harmonic of the gravitational wave signal, $f_n = n \cdot f_{\rm orb}$ is the frequency of the $n^{\rm th}$ harmonic of the gravitational wave signal, $f_{\rm orb}$ is the orbital frequency, $S_{\rm n}(f_n)$ is the LISA sensitivity curve at frequency $f_n$ \citep[e.g.][]{Robson+2019} and $h_{c,n}$ is the characteristic strain of the $n^{\rm th}$ harmonic, given by
\begin{equation}\label{eq:charstrain}
    h^2_{c,n} = \frac{2^{5/3}}{3 \pi^{4/3}} \frac{(G \mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{n f_{\rm orb}^{1/3}} \frac{g(n,e)}{F(e)}
\end{equation}
\citep[e.g.][]{Barack+2004}, where $D_L$ is the luminosity distance to the source, $f_{\rm orb}$ is the orbital frequency, $g(n, e)$ and $F(e)$ are given in \citet{Peters+1963} and $\mathcal{M}_c$ is the chirp mass, defined as
\begin{equation}\label{eq:chirp_mass}
    \mathcal{M}_c = \frac{(m_1 m_2)^{3/5}}{(m_1 + m_2)^{1/5}}.
\end{equation}

For the sake of computational efficiency, it is convenient to make the approximation that many of the binaries are stationary in frequency space on the timescale of the observing mission. This simplifies the signal-to-ratio equation to
\begin{equation}\label{eq:snr_stat}
    \rho_{\rm stat}^2 = \sum_{n=1}^{\infty} \frac{2 \, T_{\rm obs} h_n^2}{S_n(f_n)},
\end{equation}
where $T_{\rm obs}$ is the observing time and $h_n$ is the strain amplitude of gravitational waves emitted in the $n^{\rm th}$ harmonic, given by
\begin{equation}
    h_n^2 = \frac{2^{25/3}}{5} \frac{\left(G \mathcal{M}_{c}\right)^{10/3}}{c^{8}} \frac{\left(\pi f_{\mathrm{orb}}\right)^{4 / 3}}{D_L^2} \frac{g(n, e)}{n^2}.
\end{equation}

We consider a binary as stationary if its change in gravitational wave frequency is less than 1\% of the initial frequency. None binaries in our sample are in the non-stationary regime and therefore we believe that this is an acceptable approximation.

\subsection{Detection rate}
We calculate the detection rate for any one Milky Way instance by determining the fraction of binaries that are detectable and normalising to the true mass and IMF of the Milky Way in order to account for our relatively smaller galaxy size and restricted mass ranges. For a more in-depth discussion of this normalisation process see Appendix~\ref{app:rate_normalisation}.

>>>>

\subsubsection{Population overview}

@@

\subsubsection{Initial conditions}

@@

\subsection{Galaxy synthesis}

@@

@@

\subsection{Gravitational wave detection}
We use the Python package \href{https://legwork.readthedocs.io/en/latest/}{LEGWORK} to evolve binaries and calculate their LISA detectability. For a full derivation of the equations given below please see the LEGWORK release paper (Wagg et al. in prep) or \href{https://legwork.readthedocs.io/en/latest/notebooks/Derivations.html}{documentation}.

\subsubsection{Inspiral evolution}

@@

\subsubsection{Binary detectability}

We define a binary as detectable if its gravitational wave signal has a signal-to-noise ratio of greater than 7 \citep[e.g.][]{Breivik+2020, Korol+2020}. The sky-, polarisation- and orientation-averaged signal-to-noise ratio, $\rho$, of an inspiraling binary can be calculated with the following \citep[e.g.][]{Finn+2000}
\begin{equation}\label{eq:snr}
    \rho^2 = \sum_{n=1}^{\infty} \int_{f_{n, i}}^{f_{n, f}} \frac{h_{c, n}^{2}}{f_{n}^{2} S_{\rm n}\left(f_{n}\right)} \dd{f_n},
\end{equation}
where $n$ is a harmonic of the gravitational wave signal, $f_n = n \cdot f_{\rm orb}$ is the frequency of the $n^{\rm th}$ harmonic of the gravitational wave signal, $f_{\rm orb}$ is the orbital frequency, $S_{\rm n}(f_n)$ is the LISA sensitivity curve at frequency $f_n$ \citep[e.g.][]{Robson+2019} and $h_{c,n}$ is the characteristic strain of the $n^{\rm th}$ harmonic, given by \citep[e.g.][]{Barack+2004}
\begin{equation}\label{eq:charstrain}
    h^2_{c,n} = \frac{2^{5/3}}{3 \pi^{4/3}} \frac{(G \mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{f_{\rm orb}^{1/3}} \frac{g(n,e)}{n F(e)},
\end{equation}
where $D_L$ is the luminosity distance to the source, $f_{\rm orb}$ is the orbital frequency, $g(n, e)$ and $F(e)$ are given in \citet{Peters+1963} and $\mathcal{M}_c$ is the chirp mass, defined as
\begin{equation}\label{eq:chirp_mass}
    \mathcal{M}_c = \frac{(m_1 m_2)^{3/5}}{(m_1 + m_2)^{1/5}}.
\end{equation}

We use LEGWORK to calculate the signal-to-noise ratio for each binary and the package ensures that enough harmonics are computed for each binary such that the error on the gravitational wave luminosity remains below 1\%.

\subsubsection{Detection rate calculation}
For each physics variation model, we compute the total number of merging DCOs in the Milky Way today based on the COMPAS simulation. This takes into account the true mass and IMF of the Milky Way since our simulations each have a slightly smaller galaxy size and only sample from massive stars rather than the full IMF.

We determine the fraction of binaries that are detectable in each Milky Way instance by summing the adaptive importance sampling weights of the binaries that have an SNR greater than 7 and dividing by the total weights in the simulation. We multiply this fraction by the total number in the Milky Way to find a detection rate.
\begin{equation}
    N_{\rm detect} = \frac{\sum_{i = 0}^{N_{\rm detect}} w_i}{\sum_{i = 0}^{N_{\rm DCO}} w_i} \cdot N_{\rm MW}
\end{equation}
For a more in-depth discussion of this normalisation process see Appendix~\ref{app:rate_normalisation}.
